Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-11-04 forecast/paper/paper.tex
minor data deposit revision issues

\dropcap{T}racking and forecasting indicators from public health reporting
streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
crucial for understanding disease spread, correctly formulating public policy
responses, and rationally planning future public health resource needs.  A
companion paper \cite{Reinhart:2021} describes our research group's efforts,
beginning in April 2020, in curating, and maintaining a database of real-time
indicators that track COVID-19 activity and other relevant phenomena. The
signals (a term we use synonomously with ``indicators'') in this database are
accessible through the COVIDcast API \cite{CovidcastAPI}, as well as associated
R \cite{CovidcastR} and Python \cite{CovidcastPy} packages, for convenient data  
fetching and processing. In the current paper, we quantify the utility provided
by a core set of these indicators for two fundamental prediction tasks:
probabilistic forecasting of COVID-19 case rates and prediction of future
COVID-19 case hotspots (defined by the event that a relative increase in
COVID-19 cases exceeds a certain threshold).

At the outset, we should be clear that our intent in this paper is \textit{not}
to provide an authoritative take on cutting-edge COVID-19 forecasting methods.
Similarly, some authors, e.g., \cite{ioannidis2020a}, have pointed out numerous
mishaps of forecasting during the pandemic and it is not our general intent to
fix them here. Instead, we start with a basic and yet reasonably effective
predictive model for predicting future trends in COVID-19 cases, and present a
rigorous, quantitative assessment of the added value provided by auxiliary
indicators, that are derived from data sources that operate outside of 
traditional public health streams. In particular, we consider five indicators
derived from de-identified medical insurance claims, self-reported symptoms from
online surveys, and COVID-related Google searches.

@@

As for geographic resolution, we consider the prediction of COVID-19 case rates
and hotspots aggregated at the level of an individual \textit{hospital referral
  region} (HRR). HRRs correspond to groups of counties in the United States
within the same hospital referral system. The Dartmouth Atlas of Healthcare 
Policy~\cite{DartmouthHRR}, defines these 306 regions based on a number of
characteristics. They are contiguous regions such that most of the hospital
services for the underlying population are performed by hospitals within the
region. Each HRR also contains at least one city where major procedures
(cardiovascular or neurological) are performed. The smallest HRR has a
population of about 125,000. While some are quite large (such as the one
containing Los Angeles, which has more than 10 million people), generally HRRs  
are much more homogenous in size than the (approximately) 3200 U.S. counties,   
and they serve as a nice middle ground in between counties and states.

@@

Below is summary of the high-level conclusions.

>>>>

\dropcap{T}racking and forecasting indicators from public health reporting
streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
crucial for understanding disease spread, correctly formulating public policy
responses, and rationally planning future public health resource needs.  A
companion paper \cite{Reinhart:2021} describes our research group's efforts,
beginning in April 2020, in curating and maintaining a database of real-time
indicators that track COVID-19 activity and other relevant phenomena. The
signals (a term we use synonomously with ``indicators'') in this database are
accessible through the COVIDcast API \cite{CovidcastAPI}, as well as associated
R \cite{CovidcastR} and Python \cite{CovidcastPy} packages, for convenient data  
fetching and processing. In the current paper, we quantify the utility provided
by a core set of these indicators for two fundamental prediction tasks:
probabilistic forecasting of COVID-19 case rates and prediction of future
COVID-19 case hotspots (defined by the event that a relative increase in
COVID-19 cases exceeds a certain threshold).

At the outset, we should be clear that our intent in this paper is \textit{not}
to provide an authoritative take on cutting-edge COVID-19 forecasting methods.
Similarly, some authors, e.g., \cite{ioannidis2020a}, have pointed out numerous
mishaps of forecasting during the pandemic, and it is not our general intent to
fix them here. Instead, we start with a basic and yet reasonably effective
predictive model for predicting future trends in COVID-19 cases, and present a
rigorous, quantitative assessment of the added value provided by auxiliary
indicators, that are derived from data sources that operate outside of 
traditional public health streams. In particular, we consider five indicators
derived from de-identified medical insurance claims, self-reported symptoms from
online surveys, and COVID-related Google searches.

@@

As for geographic resolution, we consider the prediction of COVID-19 case rates
and hotspots aggregated at the level of an individual \textit{hospital referral
  region} (HRR). HRRs correspond to groups of counties in the United States
within the same hospital referral system. The Dartmouth Atlas of Healthcare 
Policy~\cite{DartmouthHRR}, defines these 306 regions based on a number of
characteristics. They are contiguous regions such that most of the hospital
services for the underlying population are performed by hospitals within the
region. Each HRR also contains at least one city where major procedures
(cardiovascular or neurological) are performed. The smallest HRR has a
population of about 125,000. While some are quite large (such as the one
containing Los Angeles, which has more than 10 million people), generally HRRs  
are much more homogenous in size than the (approximately) 3200 U.S.\ counties,   
and they serve as a nice middle ground in between counties and states.

@@

Below is a summary of the high-level conclusions.
